A Gentle Introduction and Application of Feature-Based Clustering with Psychological Time Series

### 2023-08-11 sections/03FeatureExtraction.Rmd
typo fixed

\paragraph{Self-similarity.}
Self-similarity in time series data refers to the property of a time series to exhibit similar patterns of behavior over different time scales \citep{dmello2021}. That is, self-similarity describes how much a measurement carries over to future measurements. One important self-similarity in psychological time series is \texit{inertia} --- how much a measurement carries over to its next measurement \citep{kuppens2010, suls1998}. If inertia is high a development tends to stay in a certain state. Because high inertia is resistant to change, in emotion dynamics high inertia of negative affect has been found to be indicative of under-reactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}. In a similar vein, high inertia in negative affect at baseline was even predictive of the initial onset of depression \citep{kuppens2012}. Conceptually, inertia is more broadly connected to research questions such as: ``Do patients stay in a negative mood for several measurements?'' or ``Do migrants stay with their language practice for several days at a time?''

>>>>

\paragraph{Self-similarity.}
Self-similarity in time series data refers to the property of a time series to exhibit similar patterns of behavior over different time scales \citep{dmello2021}. That is, self-similarity describes how much a measurement carries over to future measurements. One important self-similarity in psychological time series is \textit{inertia} --- how much a measurement carries over to its next measurement \citep{kuppens2010, suls1998}. If inertia is high a development tends to stay in a certain state. Because high inertia is resistant to change, in emotion dynamics high inertia of negative affect has been found to be indicative of under-reactive systems and to be characteristic of psychological maladjustment \citep{kuppens2010}. In a similar vein, high inertia in negative affect at baseline was even predictive of the initial onset of depression \citep{kuppens2012}. Conceptually, inertia is more broadly connected to research questions such as: ``Do patients stay in a negative mood for several measurements?'' or ``Do migrants stay with their language practice for several days at a time?''
